---
title: "conditionals_loop"
author: "Colleen McCamy"
date: "2022-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)

```


# Part 1

Task 1

```{r}

# printing a response based on conditionals and an object

pm2_5 <- 180

if (pm2_5 < 100) {
  print ("Low to moderate risk")
} else if (pm2_5 >= 100 | pm2_5 < 150) {
  print ("Unhealthy for sensitive groups")
} else if (pm2_5 >= 150){print("Health risk present")
  }

```

Task 2

```{r}

# using string detect to print an output based on vector values

species <- c("blue fish")

if (str_detect(species, pattern = "whale")) {
  print("I found a whale!")
}

```


Task 3

```{r}
# using a vector to determine an output via if else statement

max_airtemp_c <- c(24.1)

if (max_airtemp_c > 27){
  print ("Temperature too high")
} else print ("Temperature OK")

```

Task 4

```{r}

base_burrito <- 6.50

main_ingredient <- c("chicken")

if (main_ingredient == "veggie"){
  print(base_burrito)
}else if (main_ingredient == "chicken"){
  print(base_burrito + 3.00)
}else if (main_ingredient == "steak"){print(base_burrito + 3.25)}


```

Task 5:
Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.

```{r}

fish <- c(8, 10, 12, 23)
fish_types <- c("goldfish", "tetras", "guppies", "mollies")

for (i in seq_along(fish_types)) {
  if (fish_types[i] == "goldfish"){
  fish_prop <- fish[i] / sum(fish)
} else if (fish_types[i] == "tetras"){
  fish_prop <- fish[i] / sum(fish)
} else if (fish_types[i] == "guppies"){
  fish_prop <- fish[i] / sum(fish)
} else if (fish_types[i] == "mollies"){
  fish_prop <- fish[i] / sum(fish)
}
}


```

Task 6
There is an existing vector in R called month.name that contains all month names (just try running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.

Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).

```{r}


for (i in seq_along(month.name)){
  print(paste(month.name[i], "is month", i))
}


# index based on i and then directly pull i
```

